### 2022-05-17 v1/README.Rmd
Version 2.0.0 initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%"
)
library(badger)
```

<img src="man/figures/logo.png" width="200" align="right"/>

# Fasano-Franceschini Test

<!-- badges: start -->

`r badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
[![R build status](https://github.com/nesscoder/fasano.franceschini.test/workflows/R-CMD-check/badge.svg)](https://github.com/nesscoder/fasano.franceschini.test/actions) 
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
`r badge_devel("nesscoder/fasano.franceschini.test", "blue")` 
`r badge_code_size("nesscoder/fasano.franceschini.test")`
`r badge_cran_download("fasano.franceschini.test", "grand-total", "blue")`

<!-- badges: end -->

The `fasano.franceschini.test` package is an R implementation of the 2-D Kolmogorov-Smirnov (**KS**) two-sample test as defined by Fasano and Franceschini (1987). This is a variant of the 2-D two-sample KS test as originally defined by Peacock (1983).

## Installation

You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("fasano.franceschini.test")
```

And the development version of the `fasano.franceschini.test` package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nesscoder/fasano.franceschini.test")
```

## Example

#### Underlying Distributions Are Different

```{r}
library(fasano.franceschini.test)

#set seed for reproducible example
set.seed(123)

#create 2-D samples with the different underlying distributions
sample1Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 3),
  y = rnorm(n = 50, mean = 0, sd = 1)
)
sample2Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 1),
  y = rnorm(n = 50, mean = 0, sd = 3)
)

fasano.franceschini.test(sample1Data,sample2Data)
```

#### Underlying Distributions Are The Same

```{r}
#set seed for reproducible example
set.seed(123)

#create 2-D samples with the same underlying distributions
sample1Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 1),
  y = rnorm(n = 50, mean = 0, sd = 1)
)
sample2Data <- data.frame(
  x = rnorm(n = 50, mean = 0, sd = 1),
  y = rnorm(n = 50, mean = 0, sd = 1)
)

fasano.franceschini.test(sample1Data,sample2Data)
```