### 2023-05-04 README.Rmd
Updated docs

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%",
  warning = TRUE,
  message = TRUE
)
library(badger)
```

<img src="man/figures/logo.png" width="200" align="right"/>

# Fasano-Franceschini Test

<!-- badges: start -->

`r badge_custom(x = "arXiv", y = "abs/2106.10539", color = "yellow", url = "https://arxiv.org/abs/2106.10539")` 
`r badge_cran_release("fasano.franceschini.test", "orange")`
`r badge_devel("nesscoder/fasano.franceschini.test", "blue")` 
`r badge_cran_download("fasano.franceschini.test", "grand-total", "green")`
`r badge_github_actions("nesscoder/fasano.franceschini.test")`
<!-- badges: end -->

The `fasano.franceschini.test` package is an R implementation of the multidimensional Kolmogorov-Smirnov two-sample test as defined by Fasano and Franceschini (1987).

```
Fasano, G. & Franceschini, A. (1987). A multidimensional version of the
Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society,
225:155-170. doi: 10.1093/mnras/225.1.155.
```

## Installation

You can install the released version of the `fasano.franceschini.test` package from [CRAN](https://cran.r-project.org/web/packages/fasano.franceschini.test/index.html) with:

``` r
install.packages("fasano.franceschini.test")
```

The development version of the `fasano.franceschini.test` package can be installed from [GitHub](https://github.com/braunlab-nu/fasano.franceschini.test/) with:

``` r
# install.packages("devtools")
devtools::install_github("nesscoder/fasano.franceschini.test")
```

## Examples

#### Underlying distributions are the same

```{r}
library(fasano.franceschini.test)

# set seed for reproducibility
set.seed(0)

# create 2D samples from the same underlying distribution
S1 <- data.frame(x = rnorm(n = 50, mean = 0, sd = 1),
                 y = rnorm(n = 50, mean = 0, sd = 3))
S2 <- data.frame(x = rnorm(n = 100, mean = 0, sd = 1),
                 y = rnorm(n = 100, mean = 0, sd = 3))

fasano.franceschini.test(S1, S2, seed = 0)
```

#### Underlying distributions are different

```{r}
# set seed for reproducibility
set.seed(1)

# create 3D samples from different underlying distributions
S1 <- cbind(rgamma(n = 43, shape = 2),
            rpois(n = 43, lambda = 5),
            rpois(n = 43, lambda = 3.5))
S2 <- cbind(rgamma(n = 72, shape = 2),
            rpois(n = 72, lambda = 5),
            rpois(n = 72, lambda = 5))

fasano.franceschini.test(S1, S2, seed = 1)
```