### 2024-02-15 README.Rmd
2.2.3

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/",
  out.width = "100%",
  warning = TRUE,
  message = TRUE
)
library(badger)
```

<img src="man/figures/logo.png" width="200" align="right"/>

# Fasano-Franceschini Test

<!-- badges: start -->

`r badge_custom(x = "R Journal", y = "10.32614/RJ-2023-067", color = "yellow", url = "https://doi.org/10.32614/RJ-2023-067")`
`r badge_cran_release("fasano.franceschini.test", "orange")`
`r badge_devel("braunlab-nu/fasano.franceschini.test", "blue")` 
`r badge_cran_download("fasano.franceschini.test", "grand-total", "green")`
`r badge_github_actions("braunlab-nu/fasano.franceschini.test")`
<!-- badges: end -->

The `fasano.franceschini.test` package is an R implementation of the multivariate Kolmogorov-Smirnov two-sample test as defined by Fasano and Franceschini (1987).
```
Fasano, G. & Franceschini, A. (1987). A multidimensional version of the
Kolmogorov-Smirnov test. Monthly Notices of the Royal Astronomical Society,
225:155-170. doi: 10.1093/mnras/225.1.155.
```
A manuscript accompanying this package was published in The R Journal.
```
Puritz, C., Ness-Cohn, E., Braun, R. (2023). fasano.franceschini.test: An Implementation
of a Multivariate KS Test in R. The R Journal, 15(3):159-171. doi: 10.32614/RJ-2023-067.
```

## Installation

You can install the released version of the `fasano.franceschini.test` package from CRAN with:
``` r
install.packages("fasano.franceschini.test")
```

The development version of the `fasano.franceschini.test` package can be installed from GitHub with:
``` r
if (!require("devtools", quietly = TRUE))
    install.packages("devtools")
devtools::install_github("braunlab-nu/fasano.franceschini.test")
```

## Examples

#### Underlying distributions are the same
```{r}
library(fasano.franceschini.test)

# Set seed to draw reproducible samples
set.seed(0)

# Create 2D samples from the same underlying distribution
S1 <- data.frame(x = rnorm(n = 50, mean = 0, sd = 1),
                 y = rnorm(n = 50, mean = 0, sd = 3))
S2 <- data.frame(x = rnorm(n = 100, mean = 0, sd = 1),
                 y = rnorm(n = 100, mean = 0, sd = 3))

# Set seed to generate reproducible p-values
fasano.franceschini.test(S1, S2, seed = 0)
```

#### Underlying distributions are different
```{r}
# Set seed to draw reproducible samples
set.seed(1)

# Create 3D samples from different underlying distributions
S1 <- cbind(rgamma(n = 43, shape = 2),
            rpois(n = 43, lambda = 5),
            rpois(n = 43, lambda = 3.5))
S2 <- cbind(rgamma(n = 72, shape = 2),
            rpois(n = 72, lambda = 5),
            rpois(n = 72, lambda = 5))

# Set seed to generate reproducible p-values
fasano.franceschini.test(S1, S2, seed = 1)
```